---
title: "Political Leaning in Correlation with COVID-19 Outcomes"
author: "Ben Leland, Krithik Duraisamy, Selma Link, Valerie Lam"
date: "11/13/22"
output: html_document
---
```{r setup, include = FALSE}
source("../source/index_data.R")
```
# Summary
The use of this report is to inform those of the correlation between political leaning and containment and neutralization of the COVID-19 virus. Since its discovery in late 2019, COVID-19 has been a growing issue and evaluating the success of efforts to contain it is an important issue. To identify similarities between politics and neutralization efforts multiple points of data will be evaluated. The data being: election results, vaccine, hesitancy, death counts, and mask use. To begin, the election held in 2020 contained vital information for the political leanings of the public. The votes are as such: Donald Trump had `r num_votes_trump()` votes, whereas Joe Biden had `r num_votes_biden()` votes. This states that majority of the US population was leaning towards democrat and higher restrictions on protection against COVID-19. From this, the average percentage of people who actually wore masks can be correlated to political leaning. `r never_mask_avg` percent of people did not choose to wear a mask outdoors, `r sometimes_mask_avg` percent of people sometimes chose to wear a mask, and `r always_mask_avg` percent of people always chose to wear a mask. From this data it can be deduced that 51% of the population chose to wear a mask outdoors correlating to the data that democrats were majority of that population at the time. In contrast, those who chose to not wear masks as often fell ill quicker and easier. The average number of deaths per county in the US from January 1, 2020 to October 22, 2022 was `r avg_covid_deaths`. An idea can be construed that death counts can be correlated to mask usage further proven by political leaning and policies made between 2020 and 2022. Further evidence is needed to prove this and will be produced in visual displays.

### Table
```{r table, echo = FALSE, include = TRUE}
# Here display the code to produce the table
# Please change or modify the code chunk to produce the desired outcome

```
Summary

### Chart 1
```{r chart 1, echo = FALSE, include = TRUE}
# Here display the code to produce chart 1
# Please change or modify the code chunk to produce the desired outcome

```
Summary

### Chart 2
```{r chart 2, echo = FALSE, include = TRUE}
# Here display the code to produce chart 2
# Please change or modify the code chunk to produce the desired outcome

```
Summary

### Chart 3
```{r chart 3, echo = FALSE, include = TRUE}
# Here display the code to produce chart 3
# Please change or modify the code chunk to produce the desired outcome

```
Summary
